Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming

### 2017-02-02 gassolarpaper/gassolarpaper.tex
shorten captions

$b$ \> wing span [ft] \\
BSFC \> brake specific fuel consumption [kg/kW/hr] \\
$\text{BSFC}_{100\%}$ \> BSFC at full throttle [kg/kW/hr] \\
$c$ \> wing chord [m] \\

@@

$d$ \> tail boom diameter [m] \\
$D_{\text{boom}}$ \> tail boom drag [N] \\
$\Delta W$ \> wing section weight [N] \\
$\Delta y$ \> wing section length [m] \\

@@

$E$ \> Young's Modulus [Pa] \\
$E_{\text{batt}}$ \> energy stored in battery [J] \\
$E_{\text{sun}}$ \> total solar energy available [Whr] \\
$(E/S)_{\text{twilight}}$ \> twilight solar energy [Whr/m$^2$] \\
$(E/S)_{\text{day}}$ \> daily required solar energy [Whr/m$^2$] \\
$(E/S)_{\text{sun}}$ \> available solar energy [Whr/m$^2$] \\

@@

$g$ \> gravitational constant [m/s$^2$] \\
$h$ \> flight altitude [ft] \\
$h_{\text{batt}}$ \> battery specific energy [Whr/kg] \\
$h_{\text{cap}}$ \> spar cap separation [m] \\
$I$ \> cap spar moment of inertia [m$^4$] \\
$I_0$ \> tail boom root moment of inertia [m$^4$] \\

@@

$K_q$ \> wing loading constant [N/m$^2$] \\
$L$ \> atmospheric lapse rate [K/m] \\
$L_\text{h}$ \> horizontal tail lift [N] \\
$l_\text{fuse}$ \> fuselage length [m] \\
$l_\text{h}$ \> horizontal tail moment arm [m] \\
$l_\text{v}$ \> vertical tail moment arm [m] \\
$m$ \> tail boom mass [kg] \\
$m'$ \> local wing mass [kg/m] \\

@@

$\dot{m}_{\text{fuel}}$ \> fuel flow rate [kg/s] \\
$\mathcal{M}$ \> wing bending moment [Nm] \\
MTOW \> max take-off weight [N] \\

@@

$p$ \> air pressure [Pa]\\
$p_{11}$ \> air pressure at 11,000 meters [Pa]\\

@@

$P_{0}$ \> magnitude of available solar power [W/m$^2$] \\
$P_{\text{avionics}}$ \> avionics power [hp] \\
$P_{\text{oper}}$ \> aircraft operating power [hp] \\
$(P/S)_{\text{min}}$ \> minium operational solar power [Whr/m$^2$] \\
$P_{\text{shaft}}$ \> engine/motor shaft power [hp] \\
$P_{\text{sun}}$ \> useable solar power per unit area [W/m$^2$] \\
$P_{\text{sun surface}}$ \> power emitted at the sun's surface [W/m$^2$] \\

@@

$R$ \> aircraft range [nmi] \\

@@

$R_{\text{fuse}}$ \> fuselage radius [m] \\
$R_{\text{spec}}$ \> specific gas constant [J/k/K] \\

@@

$q$ \> distributed wing loading [N/m] \\

@@

$S$ \> wing planform area [m$^2$]\\
$\mathcal{S}$ \> shear forces [N] \\
$S_{\text{fuse}}$ \> fuselage wetted surface area [m$^2$]\\
$S_{\text{h}}$ \> horizontal tail planform area [m$^2$]\\
$S_{\text{solar}}$ \> solar cell area [m$^2$]\\
$S_{\text{v}}$ \> vertical tail planform area [m$^2$]\\
$t$ \> flight time [days] \\
$t_0$ \> tail boom root wall thickness [m] \\
$t_{\text{cap}}$ \> spar cap thickness [m] \\
$t_{\text{day}}$ \> hours of daylight [hr] \\
$t_{\text{night}}$ \> hours of darkness [hr] \\
$t_{\text{sun rise}}$ \> time of sun rise [hr] \\
$t_{\text{sun set}}$ \> time of sun set [hr] \\
$T$ \> aircraft thrust [N] \\
$T_{11}$ \> air temperature at 11,000 meters [K] \\
$T_{\text{atm}}$ \> air temperature at flight altitude [K] \\
$T_{\text{sl}}$ \> air temperature at sea level [K] \\
$V$ \> true airspeed [m/s] \\
$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume [m$^3$] \\
$V_{\text{gust}}$ \>  gust velocity [m/s] \\

@@

$V_{\text{ref}}$ \>  reference wind airspeed [m/s] \\
$V_{\text{wind}}$ \>  wind speed [m/s] \\

@@

$w$ \> wing deflection [m] \\
$W$ \> aircraft weight [N] \\
$W_{\text{ave}}$ \> average weight over a flight segment [N] \\
$W_{\text{batt}}$ \> battery weight [N] \\
$w_{\text{cap}}$ \> spar cap width [m] \\
$W_{\text{cent}}$ \> center of aircraft weight [N] \\
$W_{\text{empennage}}$ \> empennage weight [N] \\
$W_{\text{engine}}$ \> engine weight [N] \\
$W_{\text{final}}$ \> end of flight segment aircraft weight [N] \\
$W_{\text{fuel}}$ \> fuel weight [N] \\
$W_{\text{fuselage}}$ \> fuselage weight [N] \\
$W_{\text{h}}$ \> horizontal tail weight [N] \\
$W_{\text{initial}}$ \> start of flight segment aircraft weight [N] \\
$w_{\text{max}}$ \> maximum deflection limit [m] \\
$W_{\text{payload}}$ \> payload weight [N] \\
$W_{\text{skin}}$ \> wing skin weight [N] \\
$W_{\text{solar}}$ \> solar cell weight [N] \\
$W_{\text{spar}}$ \> wing spar weight [N] \\
$W_{\text{structural}}$ \> structural weight [N] \\
$W_{\text{v}}$ \> vertical tail weight [N] \\
$W_{\text{wing}}$ \> wing weight [N] \\
$y$ \> distance from wing root along span [m] \\

@@

$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities [rad] \\

@@

$\mu$ \> air viscosity [kg/m/s] \\
$\mu_0$ \> reference air viscosity [kg/m/s] \\

@@

$\rho$ \> air density [kg/m$^3$] \\
$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber [kg/cm$^2$] \\
$\rho_{\text{cfrp}}$ \> density of carbon fiber [kg/cm$^3$] \\
$\rho_{\text{foam}}$ \> density of foam [lbf/ft$^3$] \\
$\rho_{\text{solar}}$ \> solar cell density [kg/m$^2$] \\
$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress [Pa] \\

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
    \caption{\textbf{Variation of wind speed by month at 45 degrees latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.  Lower, middle and upper lines represent 85th, 90th and 95th percentile winds respectively.}}
	\label{f:windvsmonth}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswind.pdf}
    \caption{\textbf{Variation of wind speed with altitude at latitudes of 30, 35, and 45 degrees.  Left, middle and right lines represent 85th, 90th and 95th percentile winds respectively. Wind data includes winds during December for the years 2005-2015.}}
	\label{f:altvswind}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswind.pdf}
    \caption{\textbf{Variation of wind speed with latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.  Left, middle and right lines represent 85th, 90th and 95th percentile winds respectively. Wind data includes winds during December for the years 2005-2015.}}
	\label{f:latvswind}
	\end{center}
\end{figure}

@@

For example, an aircraft optimized for $\pm$30 degrees latitude will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitude between 20 and 30 degrees. 
Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
Additionally, because long-endurance aircraft are likely to fly over land and the majority of land mass lies in the Northern Hemisphere, only wind speed data from the Northern Hemisphere is considered.
In total, 41 wind speed constraints were generated from 20-60 degrees North latitude. (See Appendix A) 
The constraint on wind speed and air density at the 35 degree latitude is

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl35.pdf}
    \caption{\textbf{Comparison of a fitted equation of wind speed to air density and percentile wind speed~\eqref{e:windfit35}. (Log space RMS error = 0.043.) Wind speed data includes December wind speeds from 2005-2015 at the 35th degree latitude.}}
	\label{f:windfitl35}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
    \caption{\textbf{The available solar power per unit area during the winter solstice in the Northern Hemisphere at $\phi=30^{\circ}$. For a solar-electric powered aircraft to be operational for more than 24 hours, the area under the red curve must be greater than the areas shown in blue, magenta and green.}}
	\label{f:lat30}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
     \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions for $(P/S)_{\text{min}}$ at 30 degrees latitude. (Log-space RMS error = 0.007 and 0.0180)}}
 \label{f:energyapprox}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{JH01polar.pdf}
    \caption{\textbf{The posynomial fit, Equation~\eqref{e:aerodragprof}, of the \emph{JH01} airfoil drag polar is represented as solid lines at different Reynolds numbers.  The open circles are the calculated drag polars from XFOIL. (Log space RMS error~=~0.00489.)}}
	\label{f:JH01polar}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=517,natheight=408]{arctanfit.pdf}
    \caption{\textbf{Comparison of monomial approximation to Equation~\ref{e:gustalpha} for $V/V_{\text{gust}} \in [0,0.7]$. (Log space RMS error = 0.0246)}}
	\label{f:arctanfit}
	\end{center}
\end{figure}

@@

A cap spar is considered for the solar-electric and gas powered aircraft.  Both aircraft optimization models share the same governing equations.  A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.8\textwidth,natwidth=3242,natheight=1195]{capspar2.jpg}
    \caption{\textbf{Cross sectional view of a cap spar.  The two caps are made of carbon fiber, separated by a foam core. A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.}}
	\label{f:capspar}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
    \caption{\textbf{Comparison of XFOIL data for NACA airfoils to GP-compatible generated fit,~\eqref{e:taildrag}. (Log space RMS error = 0.0173)}}
	\label{f:taildragpolar}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.pdf}
    \caption{\textbf{Comparison of power law fit and engine data from University of North Dakota technical report.\cite{gasengine} (Log space RMS error = 0.34)}}
	\label{f:powervsweightfit}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.pdf}
    \caption{\textbf{Comparison of GP-compatible approximation to RCV manufacturing data for the DF70 engine.  Function is representative of engine performance for engines of similar size.  (Log-space RMS error is 0.007) }}
	\label{f:powervsweightfit}
	\end{center}
\end{figure}

@@

To compare the gas-powered and solar-electric architectures the optimization models were solved by minimizing the max take off weight across different bands of latitudes, shown in Figure~\ref{f:latvsmtowtrade}. The gas powered optimization model contains 606 free variables and the solar-electric powered optmization model contains between 165 and 585 free variables depending on latitude band requirement.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.pdf}
    \caption{\textbf{Trade study comparing the lowest possible max take off weight across a range of possible latitude band requirements. Each point is a unique solution from the optimization. Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.  The endurance requirement for the gas powered aircraft in this trade study is 7 days. The next integer latitude value for each of the solar-electric curves is infeasible for the given assumptions.}}
    \label{f:latvsmtowtrade}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.pdf}
    \caption{\textbf{Trade study of endurance and size for gas powered architecture., Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively. Max take off weight was minimized for this analysis.}}
	\label{f:spanvsendurance}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
     \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
 \end{subfigmatrix}
    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$. Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.}}
    \label{f:solarobjcomp}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper.pdf}
 \caption{\textbf{Comparison of optimum operating altitude for the solar-electric powered aircraft with and without a detailed structural model.  With only a structural fraction as the structural model the optimum point is at the minimum wind speed.  With a more detailed structural model the optimum operating altitude is not at the minimum wind speed because of variable interaction between wing weight, air density, and wind speed.}}
 \label{f:altoper}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission.pdf}
    \caption{\textbf{Circles and squares represent the optimized drag and lift coefficient values during the course of mission with and without the wind speed constraint. Solid lines are drag polars at various Reynolds numbers.  The maximum $C_L^{1.5}/C_D$ are marked by a plus sign on the polar curve.}}
 \label{f:polarmission}
	\end{center}
\end{figure}

The solar aircraft size and design depends on the assumed solar cell efficiency and battery specific energy. 
By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.

 \begin{figure}[H]
 \begin{subfigmatrix}{3}% number of columns
  \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.pdf}}
  \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.pdf}}
  \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.pdf}}
  \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.pdf}}
  \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.pdf}}
  \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.pdf}}
  \subfigure[25th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.pdf}}
  \subfigure[25th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.pdf}}
  \subfigure[25th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Matrix of contour plots of battery specific energy and solar cell efficiency vs wing span for solar-electric powered aircraft where the wing span is minimized. Each point on the graph corresponds to a unique design. Values of constants other than solar cell efficiency and battery specific energy are given throughout the text.  The infeasible region and contour shapes would change for different assumed constant values.}}
 \label{f:solarcontours}
\end{figure}

>>>>

$b$ \> wing span \\ % [ft] \\
BSFC \> brake specific fuel consumption \\ %[kg/kW/hr] \\
$\text{BSFC}_{100\%}$ \> BSFC at full throttle \\ %[kg/kW/hr] \\
$c$ \> wing chord \\ %[m] \\

@@

$d$ \> tail boom diameter \\ %[m] \\
$D_{\text{boom}}$ \> tail boom drag \\% [N] \\
$\Delta W$ \> wing section weight \\ %[N] \\
$\Delta y$ \> wing section length \\ %[m] \\

@@

$E$ \> Young's Modulus \\ %[Pa] \\
$E_{\text{batt}}$ \> energy stored in battery \\ %[J] \\
$E_{\text{sun}}$ \> total solar energy available \\ %[Whr] \\
$(E/S)_{\text{twilight}}$ \> twilight solar energy \\ %[Whr/m$^2$] \\
$(E/S)_{\text{day}}$ \> daily required solar energy \\ % [Whr/m$^2$] \\
$(E/S)_{\text{sun}}$ \> available solar energy \\ % [Whr/m$^2$] \\

@@

$g$ \> gravitational constant \\ % [m/s$^2$] \\
$h$ \> flight altitude \\ % [ft] \\
$h_{\text{batt}}$ \> battery specific energy \\ % [Whr/kg] \\
$h_{\text{cap}}$ \> spar cap separation \\ % [m] \\
$I$ \> cap spar moment of inertia \\ % [m$^4$] \\
$I_0$ \> tail boom root moment of inertia \\ % [m$^4$] \\

@@

$K_q$ \> wing loading constant \\ % [N/m$^2$] \\
$L$ \> atmospheric lapse rate \\ % [K/m] \\
$L_\text{h}$ \> horizontal tail lift \\ % [N] \\
$l_\text{fuse}$ \> fuselage length \\ % [m] \\
$l_\text{h}$ \> horizontal tail moment arm \\ % [m] \\
$l_\text{v}$ \> vertical tail moment arm \\ % [m] \\
$m$ \> tail boom mass \\ % [kg] \\
$m'$ \> local wing mass \\ % [kg/m] \\

@@

$\dot{m}_{\text{fuel}}$ \> fuel flow rate \\ % [kg/s] \\
$\mathcal{M}$ \> wing bending moment \\ % [Nm] \\
MTOW \> max take-off weight \\ % [N] \\

@@

$p$ \> air pressure \\ % [Pa]\\
$p_{11}$ \> air pressure at 11,000 meters \\ % [Pa]\\

@@

$P_{0}$ \> magnitude of available solar power \\ % [W/m$^2$] \\
$P_{\text{avionics}}$ \> avionics power \\ % [hp] \\
$P_{\text{oper}}$ \> aircraft operating power \\ % [hp] \\
$(P/S)_{\text{min}}$ \> minium operational solar power \\ % [Whr/m$^2$] \\
$P_{\text{shaft}}$ \> engine/motor shaft power \\ % [hp] \\
$P_{\text{sun}}$ \> useable solar power per unit area \\ % [W/m$^2$] \\
$P_{\text{sun surface}}$ \> power emitted at the sun's surface \\ % [W/m$^2$] \\

@@

$R$ \> aircraft range \\ % [nmi] \\

@@

$R_{\text{fuse}}$ \> fuselage radius \\ % [m] \\
$R_{\text{spec}}$ \> specific gas constant \\ % [J/k/K] \\

@@

$q$ \> distributed wing loading \\ % [N/m] \\

@@

$S$ \> wing planform area \\ % [m$^2$]\\
$\mathcal{S}$ \> shear forces \\ % [N] \\
$S_{\text{fuse}}$ \> fuselage wetted surface area \\ % [m$^2$]\\
$S_{\text{h}}$ \> horizontal tail planform area \\ % [m$^2$]\\
$S_{\text{solar}}$ \> solar cell area \\ % [m$^2$]\\
$S_{\text{v}}$ \> vertical tail planform area \\ % [m$^2$]\\
$t$ \> flight time \\ % [days] \\
$t_0$ \> tail boom root wall thickness \\ % [m] \\
$t_{\text{cap}}$ \> spar cap thickness \\ % [m] \\
$t_{\text{day}}$ \> hours of daylight \\ % [hr] \\
$t_{\text{night}}$ \> hours of darkness \\ % [hr] \\
$t_{\text{sun rise}}$ \> time of sun rise \\ % [hr] \\
$t_{\text{sun set}}$ \> time of sun set \\ % [hr] \\
$T$ \> aircraft thrust \\ % [N] \\
$T_{11}$ \> air temperature at 11,000 meters \\ % [K] \\
$T_{\text{atm}}$ \> air temperature at flight altitude \\ % [K] \\
$T_{\text{sl}}$ \> air temperature at sea level \\ % [K] \\
$V$ \> true airspeed \\ % [m/s] \\
$\mathcal{V}_{\text{fuse}}$ \>  fuselage volume \\ % [m$^3$] \\
$V_{\text{gust}}$ \>  gust velocity \\ % [m/s] \\

@@

$V_{\text{ref}}$ \>  reference wind airspeed \\ % [m/s] \\
$V_{\text{wind}}$ \>  wind speed \\ % [m/s] \\

@@

$w$ \> wing deflection \\ % [m] \\
$W$ \> aircraft weight \\ % [N] \\
$W_{\text{ave}}$ \> average weight over a flight segment \\ % [N] \\
$W_{\text{batt}}$ \> battery weight \\ % [N] \\
$w_{\text{cap}}$ \> spar cap width \\ % [m] \\
$W_{\text{cent}}$ \> center of aircraft weight \\ % [N] \\
$W_{\text{empennage}}$ \> empennage weight \\ % [N] \\
$W_{\text{engine}}$ \> engine weight \\ % [N] \\
$W_{\text{final}}$ \> end of flight segment aircraft weight \\ % [N] \\
$W_{\text{fuel}}$ \> fuel weight \\ % [N] \\
$W_{\text{fuselage}}$ \> fuselage weight \\ % [N] \\
$W_{\text{h}}$ \> horizontal tail weight \\ % [N] \\
$W_{\text{initial}}$ \> start of flight segment aircraft weight \\ % [N] \\
$w_{\text{max}}$ \> maximum deflection limit \\ % [m] \\
$W_{\text{payload}}$ \> payload weight \\ % [N] \\
$W_{\text{skin}}$ \> wing skin weight \\ % [N] \\
$W_{\text{solar}}$ \> solar cell weight \\ % [N] \\
$W_{\text{spar}}$ \> wing spar weight \\ % [N] \\
$W_{\text{structural}}$ \> structural weight \\ % [N] \\
$W_{\text{v}}$ \> vertical tail weight \\ % [N] \\
$W_{\text{wing}}$ \> wing weight \\ % [N] \\
$y$ \> distance from wing root along span \\ % [m] \\

@@

$\alpha_{\text{gust}}$ \> local angle of attack from gust velocities \\ % [rad] \\

@@

$\mu$ \> air viscosity \\ % [kg/m/s] \\
$\mu_0$ \> reference air viscosity \\ % [kg/m/s] \\

@@

$\rho$ \> air density \\ % [kg/m$^3$] \\
$\rho_{A_{\text{cfrp}}}$ \> area density of carbon fiber \\ % [kg/cm$^2$] \\
$\rho_{\text{cfrp}}$ \> density of carbon fiber \\ % [kg/cm$^3$] \\
$\rho_{\text{foam}}$ \> density of foam \\ % [lbf/ft$^3$] \\
$\rho_{\text{solar}}$ \> solar cell density \\ % [kg/m$^2$] \\
$\sigma_{\text{cfrp}}$ \> maximum carbon fiber stress \\ % [Pa] \\

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
    \caption{\textbf{Wind speeds by month at 30 degrees latitude.  Bands represent 80th, 90th and 95th percentile winds.}}
	\label{f:windvsmonth}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswind.pdf}
    \caption{\textbf{Wind speeds by altitude. Bands represent 80th, 90th and 95th percentile winds.}}
	\label{f:altvswind}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswind.pdf}
    \caption{\textbf{Wind speeds by latitude.  Bands represent 80th, 90th and 95th percentile winds. }}
	\label{f:latvswind}
	\end{center}
\end{figure}

@@

For example, an aircraft optimized for $\pm$30 degrees latitude will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitude between 20 and 30 degrees. 
Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
Additionally, because long-endurance aircraft are likely to fly over land and the majority of land mass lies in the Northern Hemisphere, only wind speed data from the Northern Hemisphere is considered.
In total, 41 wind speed constraints were generated from 20-60 degrees North latitude. (See Appendix A) 
The constraint on December wind speeds and air density at the 35 degree latitude is

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl35.pdf}
    \caption{\textbf{Posynomial approximation to wind speed data comparison at 35$^{\circ}$ latitude. (Log space RMS error = 0.043.)}}
	\label{f:windfitl35}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
    \caption{\textbf{Solar power on Dec 21st at $30^{\circ}$ North. 24-hour operationality achieved when red area exceeds blue, magenta and green areas. }}
	\label{f:lat30}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
     \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
 \label{f:energyapprox}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{JH01polar.pdf}
    \caption{\textbf{Posynomial fit (solid lines) to XFOIL data (circles) comparison.  (Log-space RMS error~=~0.00489.)}}
	\label{f:JH01polar}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=517,natheight=408]{arctanfit.pdf}
    \caption{\textbf{Monomial approximation to Equation~\ref{e:gustalpha} comparison for $V/V_{\text{gust}} \in [0,0.7]$. (Log-space RMS error = 0.0246)}}
	\label{f:arctanfit}
	\end{center}
\end{figure}

@@

A cap spar is considered for the solar-electric and gas powered aircraft.  Both aircraft optimization models share the same governing equations.  A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}. A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.8\textwidth,natwidth=3242,natheight=1195]{capspar2.jpg}
    \caption{\textbf{Cross sectional view of a cap spar.}}
	\label{f:capspar}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
    \caption{\textbf{Posynomial equation (solid lines) to XFOIL data (circles) comparison. (Log space RMS error = 0.0173)}}
	\label{f:taildragpolar}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.pdf}
    \caption{\textbf{Power law fit to University of North Dakota data\cite{gasengine} comparison. (Log space RMS error = 0.34)}}
	\label{f:powervsweightfit}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.pdf}
    \caption{\textbf{Representative engine performance function based on RCV Engine Ltd data.  (Log-space RMS error = 0.007) }}
	\label{f:powervsweightfit}
	\end{center}
\end{figure}

@@

To compare the gas-powered and solar-electric architectures the optimization models were solved by minimizing the max take off weight across different bands of latitudes. Figure~\ref{latvsmtowtrade} shows this trade study evaluated at the 80th, 90th and 95th percentile wind speeds.  
Each point on the graph is a unique solution to the optimization model for the stated assumptions and contants.
The gas powered optimization model contains 606 free variables and the solar-electric powered optmization model contains between 165 and 585 free variables depending on latitude band requirement.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.pdf}
    \caption{\textbf{Gas optimization model feasible for all latitudes. Next integer latitude value for each solar-electric curve is infeasible.}}
    \label{f:latvsmtowtrade}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.pdf}
    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
	\label{f:spanvsendurance}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
     \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
 \end{subfigmatrix}
    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$.}}
    \label{f:solarobjcomp}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper.pdf}
 \caption{\textbf{Comparison of structural fraction and detailed structural models highlights trade between wing weight and air density.}}
 \label{f:altoper}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission.pdf}
    \caption{\textbf{Wind speed constraint moves lift to drag ratio from optimum endurance ratio $\text{max}(C_L^{1.5}/C_D)$ (plus sign). Solid lines are drag polars.}}
 \label{f:polarmission}
	\end{center}
\end{figure}

The solar aircraft size and design depends on the assumed solar cell efficiency and battery specific energy. 
By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
Figure~\ref{f:solarcontours} shows a matrix contour map of the solar-electric powered aircraft for multiple solar cell efficiencies, battery energy densities, latitudes, and percentile wind speeds.
Each point in Figure~\ref{f:solarcontours} is a unique design and was solved by minizing wing span. 
Values of constants other than solar cell efficiency are given throughout the text.
The infeasible regions and contour shapes would charge for different assumed constant values.

 \begin{figure}[H]
 \begin{subfigmatrix}{3}% number of columns
  \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a80.pdf}}
  \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a85.pdf}}
  \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl35a90.pdf}}
  \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a80.pdf}}
  \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a85.pdf}}
  \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl30a90.pdf}}
  \subfigure[25th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a80.pdf}}
  \subfigure[25th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a85.pdf}}
  \subfigure[25th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=542,natheight=414]{bcontourl25a90.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Operating at higher latitudes requires both high performing batteries and solar cells.}}
 \label{f:solarcontours}
\end{figure}
